---
title: "PerAct_Exam"
author: "Zofia Radwańska"
date: "2024-12-09"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# downloading the packages 
library(pacman)
pacman::p_load(tidyverse, dplyr, car)
```

#PREPROCESSING OF THE DATA

##Editing of the csv's: participant id & time difference between taps
```{r}
#adding participant ids to the csv's

#folder path
data_folder <- "/Users/zofiaradwanska/Desktop/PerAct_Exam/Perc_Act_csv"

csv_files <- list.files(path = data_folder, pattern = "*.csv", full.names = TRUE)

#vector for participant ids
participant_ids <- paste0(seq_along(csv_files))

#loop through files to add participant id and tempos as columns
for (i in seq_along(csv_files)) {
  #reading the csv files
  df <- read.csv(csv_files[i])
  
  #adding a column with participant id
  df$participant_id <- participant_ids[i]
  
  #creating a column with times between events
  df$tempo <- c(NA, diff(df$timestamp))
  
  #saving the files
  write.csv(df, csv_files[i], row.names = FALSE)
  #print(paste("Updated and saved file:", csv_files[i]))
}

```

##Editing of the data frame
```{r}
#downloading the data

##csv files with participants' responses
df <- list.files(path = "/Users/zofiaradwanska/Desktop/PerAct_Exam/Perc_Act_csv", full.names = TRUE) %>% 
  lapply(read_csv) %>%
  bind_rows

###changing the variable type
df$event <- as.factor(df$event)
df$frequency <- as.factor(df$frequency)
df$task_type <- as.factor(df$task_type)
df$condition <- as.factor(df$condition)
df$early_trained <- as.logical(df$early_trained)
df$musical_experience <- as.logical(df$musical_experience)
df$gender <- as.factor(df$gender)
df$country <- as.factor(df$country)
df$l1 <- as.factor(df$l1)
df$participant_id <- as.factor(df$participant_id)

str(df)

##mood questionaire
bmis <- read.csv("/Users/zofiaradwanska/Desktop/PerAct_Exam/The Brief Mood Introspection Scale (BMIS).csv")
colnames(bmis)[2] <- "participant_id" #changing the column name so it matches the df

str(bmis)
###factorising the mood answers into 4-level variables
bmis <- bmis %>%
  mutate(across(c(Lively, Sad, Tired, Caring, Content, Gloomy, Jittery, Drowsy, Grouchy, Peppy, Nervous, Calm, Loving, Fed.up, Active), ~factor(., levels = c("Definitely do not feel", "Do not feel", "Slightly feel", "Definitely feel"))))

#merging the dfs into one
df <- merge(df, bmis, by = "participant_id")

#changing the variable names
df$task_type <- ifelse(df$task_type == "Metronome", "aud",
                ifelse(df$task_type == "Visual Pulsation", "vis",
                ifelse(df$task_type == "Both", "multi", df$task_type)))

df$frequency <- ifelse(df$frequency == "Slow", "slow",
                ifelse(df$frequency == "Fast", "fast", df$frequency))

df$event <- ifelse(df$event == "second_tapping_keypress", "after_stim",
            ifelse(df$event == "first_tapping_keypress", "before_stim", df$event))

#adding a condition column
df <- df %>% 
  mutate(condition = paste(task_type, frequency, sep = "_"))

#deleting the time difference for the first tap in second tapping condition (the inaccurate time recording)
df <- df %>%
  group_by(participant_id) %>% 
  mutate(
    is_first_after = event == "after_stim" & row_number() == which(event == "after_stim")[1], #marking the first after_stim tap
    tempo = ifelse(is_first_after, NA, tempo)
  ) %>%
  ungroup() %>% 
  select(-is_first_after)
```

##Data inspection
```{r}
#general summary of the data 
summary(df)

#initial data inspection (by conditions)
df_dirty_summary <- df %>%
  group_by(condition, event) %>%
  filter(!event %in% c("image_description", "pulsation")) %>% 
  summarize(
    mean_tempo = mean(tempo, na.rm = TRUE),
    median_tempo = median(tempo, na.rm = TRUE),
    min_tempo = min(tempo, na.rm = TRUE),
    max_tempo = max(tempo, na.rm = TRUE),
    sd_tempo = sd(tempo, na.rm = TRUE),
    .groups = "drop"  # removes all standard grouping 
  )

```

### Visual inspection
```{r}

##before_stim tempos
a <- df %>% 
  filter(event %in% "before_stim") 

boxplot(tempo ~ condition, data = a) 
boxplot(tempo ~ participant_id, data = a) 

hist(a$tempo, breaks = 1000)
plot(density(a$tempo, na.rm = TRUE))

shapiro.test(a$tempo) #data is not normally distributed (p<0.05)
qqnorm(a$tempo)

##after_stim tempos
b <- df %>% 
  filter(event %in% "after_stim") 

boxplot(tempo ~ condition, data = b) 
boxplot(tempo ~ participant_id, data = b) 

hist(b$tempo, breaks = 1000)
plot(density(b$tempo, na.rm = TRUE))

shapiro.test(b$tempo) #data is not normally distributed (p<0.05)
qqnorm(b$tempo)

## big difference in distribution of means and medians between before and after stimulus tempos
## the tempos after stimulus are much more distributed -> some effect of conditions

df_plot <- df %>%  #df for plotting (only the tapping tempos are included)
  filter(!event %in% c("pulsation", "image_description"))

#tempo differences against timeline for each condition
ggplot(df_plot, aes(timestamp, tempo, color = condition)) +
  geom_point() +
  geom_line()

#time difference against the timeline for participants
ggplot(df_plot, aes(timestamp, tempo , color = participant_id)) +
  geom_point() +
  geom_line()

#time difference against the timeline for each participant (separately)
participant_plots <- lapply(unique(df_plot$participant_id), function(id) {
  ggplot(df_plot %>% filter(participant_id == id),
         aes(x = timestamp, y = tempo, color = as.factor(participant_id))) +
    geom_point() +
    geom_line() +
    theme_minimal() +
    labs(
      title = paste("Participant", id),
      x = "timestamp",
      y = "Time Between Taps"
    ) +
    theme(legend.position = "none")
})

##individual plots
participant_plots[[2]]  # checking the plot by participant id


```

##Data cleaning
```{r}
#removal of participant 29 -> didn't understand the instructions
#removal of participant 35 ->lack of questionnaire responses
data <- df %>% 
  filter(!participant_id %in% c(29, 35))

#summary of the time differences
data_sum <- data %>% 
  group_by(participant_id, event, condition) %>% 
  filter(!event %in% c("image_description")) %>% 
  summarize(
    mean_tempo = mean(tempo, na.rm = TRUE),
    median_tempo = median(tempo, na.rm = TRUE),
    min_tempo = min(tempo, na.rm = TRUE),
    max_tempo = max(tempo, na.rm = TRUE),
    sd_tempo = sd(tempo, na.rm = TRUE),
    .groups = "drop"  # removes all standard grouping 
  )

#identification of extreme values (outliers)
outliers <- data %>%
  group_by(participant_id, event) %>%
  filter(event %in% c("before_stim", "after_stim")) %>% 
  filter(tempo < quantile(tempo, 0.25, na.rm = TRUE) - 1.5 * IQR(tempo, na.rm = TRUE) |
         tempo > quantile(tempo, 0.75, na.rm = TRUE) + 1.5 * IQR(tempo, na.rm = TRUE))
print(outliers)

#removal of outliers
cleaned_data <- data %>%
  anti_join(outliers, by = c("participant_id", "event", "tempo"))
```
### Inspection of clean data
```{r}
df_clean_summary <- cleaned_data %>%
  group_by(condition, event) %>%
  filter(!event %in% c("image_description", "pulsation")) %>% 
  summarize(
    mean_tempo = mean(tempo, na.rm = TRUE),
    median_tempo = median(tempo, na.rm = TRUE),
    min_tempo = min(tempo, na.rm = TRUE),
    max_tempo = max(tempo, na.rm = TRUE),
    sd_tempo = sd(tempo, na.rm = TRUE),
    .groups = "drop"  # removes all standard grouping 
  )
```

### Visualisation of clean data
```{r}
##before_stim clean data
a_clean <- cleaned_data %>% 
  filter(event %in% "before_stim") 

boxplot(tempo ~ condition, data = a_clean) 
boxplot(tempo ~ participant_id, data = a_clean) 

hist(a_clean$tempo, breaks = 1000)
plot(density(a_clean$tempo, na.rm = TRUE))

shapiro.test(a_clean$tempo) #data is not normally distributed (p<0.05)
qqnorm(a_clean$tempo)

##after_stim clean data
b_clean <- cleaned_data %>% 
  filter(event %in% "after_stim") 

boxplot(tempo ~ condition, data = b_clean) 
boxplot(tempo ~ participant_id, data = b_clean) 

hist(b_clean$tempo, breaks = 1000)
plot(density(b_clean$tempo, na.rm = TRUE))

shapiro.test(b_clean$tempo) #data is not normally distributed (p<0.05)
qqnorm(b_clean$tempo)

clean_df_plot <- cleaned_data %>% 
  filter(event %in% c("before_stim", "after_stim"))

#tempo differences against timeline for each condition
ggplot(clean_df_plot, aes(timestamp, tempo, color = condition)) +
  geom_point() +
  geom_line()

#time dcleaned_data#time difference against the timeline for participants
ggplot(clean_df_plot, aes(timestamp, tempo , color = participant_id)) +
  geom_point() +
  geom_line()

clean_participant_plots <- lapply(unique(clean_df_plot$participant_id), function(id) {
  ggplot(df_plot %>% filter(participant_id == id),
         aes(x = timestamp, y = tempo, color = as.factor(participant_id))) +
    geom_point() +
    geom_line() +
    theme_minimal() +
    labs(
      title = paste("Participant", id),
      x = "timestamp",
      y = "Time Between Taps"
    ) +
    theme(legend.position = "none")
})

##individual plots
clean_participant_plots[[2]]  # checking the plot by participant id

#num 2 has some fucked up data point

```

##Participant statistics 
```{r}
#mean age
mean_age = mean(df$participant_age)
print(paste("Mean age:", mean_age))

#sd age
sd_age = sd(df$participant_age)
print(paste("sd age:", sd_age))

#total number of languages 
nr_lang = length(unique(df$l1))
print(paste("Number of L1:", nr_lang))

#total number of native countries
native_country = length(unique(df$country))
print(paste("Number of nationalities:", native_country))

#new df with only ID and l1
df_w_lang <- df %>% group_by(participant_id) %>% summarise(l1 = first(l1)) 
print(df_w_lang)

#counting how many speakers there are of each language
nr_speaker <- df_w_lang %>% group_by (l1) %>% summarise(count = n())
nr_speaker

#separating those with musical training 
df_music <- df %>% filter (musical_experience == "Yes") %>% mutate(formal_education = as.numeric(formal_education))

#finding out how many had musical training
part_music <- length(unique(df_music$participant_id))
part_music

#mean of years of training 
mean_training = mean(df_music$formal_education, na.rm = TRUE)
mean_training
print(paste("Mean years of training", mean_training))

sd_train = sd(df_music$formal_education, na.rm = TRUE)
sd_train

#separating gender
df_gend <- df %>% group_by(participant_id) %>% summarise(gender = first(gender))

#counting genders 
nr_gend <- df_gend %>% group_by(gender) %>% summarise(count = n())
nr_gend
  
#new df with only ID and l1
df_w_lang <- df %>% group_by(participant_id) %>% summarise(l1 = first(l1)) 
print(df_w_lang)

#counting how many speakers there are of each language
nr_speaker <- df_w_lang %>% group_by (l1) %>% summarise(count = n())
nr_speaker
```

#DATA ANALYSIS

##Extraction of participants with abnormal tempos (outside of expected normal tempo range)
```{r}
#tempo comparison for before_stim and pulsation events
tempo_comp <- cleaned_data %>% 
  filter(!event %in% "image_description") %>% 
  group_by(participant_id, frequency, event) %>% 
  summarise(mean_tempo = mean(tempo, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = event, values_from = mean_tempo) %>% 
  group_by(frequency) %>%
  #in case pulsation tempo was not recorded the mean for pulsation tempo within the frequency is used to compare
  mutate(
    pulsation = if_else(
      is.na(pulsation),
      mean(pulsation, na.rm = TRUE),
      pulsation
    )
  ) %>%
  ungroup()

pulsation_mean <- tempo_comp %>% 
  group_by(frequency) %>% 
  summarise(mean_pulsation = mean(pulsation, na.rm = TRUE))

#creating a column for classification of the tempo comparisons
tempo_comp <- tempo_comp %>% 
  mutate(comparison = ifelse(before_stim > pulsation, "slower", "faster"))

#merging the dfs to add the column to primary cleaned df
cleaned_data <- cleaned_data %>% 
  left_join(tempo_comp %>% select(participant_id, frequency, comparison),
            by = c("participant_id", "frequency"))

##slower than slow: 18 & 27
##faster than fast: 2
```

##Calculating the change between pre-stimulus and post_stimulus tempos
```{r}
#normalised tempo change between pre and post stimulus tapping (MEDIAN)
normalised_tempo_median <- cleaned_data %>%
  filter(event %in% c("before_stim", "after_stim")) %>%
  group_by(participant_id, event, task_type, frequency, condition, comparison) %>%
  summarise(median_tempo = median(tempo, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = event, values_from = c(median_tempo), names_prefix = "tempo_") %>%
  mutate(median_difference = tempo_after_stim - tempo_before_stim, median_normalised_difference = (tempo_after_stim - tempo_before_stim) / tempo_before_stim)

#normalised tempo change between pre and post stimulus tapping (MEAN)
normalised_tempo_mean <- cleaned_data %>%
  filter(event %in% c("before_stim", "after_stim")) %>%
  group_by(participant_id, event, task_type, frequency, condition, comparison) %>%
  summarise(mean_tempo = mean(tempo, na.rm = TRUE), .groups = "drop") %>%
  pivot_wider(names_from = event, values_from = c(mean_tempo), names_prefix = "tempo_") %>%
  mutate(mean_difference = tempo_after_stim - tempo_before_stim, mean_normalised_difference = (tempo_after_stim - tempo_before_stim) / tempo_before_stim)


normalised_tempo <- normalised_tempo_median %>% 
  left_join(normalised_tempo_mean %>% select(participant_id, task_type, frequency, condition, comparison, mean_difference, mean_normalised_difference),
            by = c("participant_id", "task_type", "frequency", "condition", "comparison"))

# responses with a "-" -> participant became faster
# responses without a "-" -> participant became slower
```

###Testing for normal distribution of tempo differences
```{r}
# tempo median difference -> normal distribution
shapiro.test(normalised_tempo$median_normalised_difference) #p-value = 0.08488 -> normal distribution
shapiro.test(normalised_tempo$median_difference) #p-value = 0.07382 -> normal distribution

# tempo mean difference -> non normal distribution
shapiro.test(normalised_tempo$mean_normalised_difference) #p-value = 0.03619 -> non normal distribution
shapiro.test(normalised_tempo$mean_difference) #p-value = 0.01267 -> non normal distribution

## regardless of the results median probably better to use -> skewed data distribution
```

##H1: The auditory stimuli will influence the spontaneous motor tempo more than the visual stimuli. 

###Assumption testing: two-way ANOVA -> all conditions are normally distributed and have homogenous variance
```{r}
#distribution of tempo differences within each conditon

frequency_diff_distribution <- normalised_tempo %>% 
  group_by(frequency) %>% 
  summarise(shapiro_test = list(shapiro.test(median_normalised_difference)), .groups = "drop") %>% 
  mutate(
    statistic = map_dbl(shapiro_test, ~ .x$statistic),
    p_value = map_dbl(shapiro_test, ~ .x$p.value)
  )
## fast: p value = 0.1069100	-> normal distribution
## slow: p value = 0.1922093	-> normal distribution

comparison_diff_distribution <- normalised_tempo %>% 
  group_by(comparison) %>% 
  summarise(shapiro_test = list(shapiro.test(median_normalised_difference)), .groups = "drop") %>% 
  mutate(
    statistic = map_dbl(shapiro_test, ~ .x$statistic),
    p_value = map_dbl(shapiro_test, ~ .x$p.value)
  )
## faster: p value = 0.2832534	-> normal distribution
## slower: p value = 0.2258640	-> normal distribution

task_type_diff_distribution <- normalised_tempo %>% 
  group_by(task_type) %>% 
  summarise(shapiro_test = list(shapiro.test(median_normalised_difference)), .groups = "drop") %>% 
  mutate(
    statistic = map_dbl(shapiro_test, ~ .x$statistic),
    p_value = map_dbl(shapiro_test, ~ .x$p.value)
  )
## aud: p value = 0.09689013 -> normal distribution
## vis: p value = 0.51504719 -> normal distribution
## multi: p value = 0.89202689 -> normal distribution

condition_diff_distribution <- normalised_tempo %>% 
  group_by(condition) %>% 
  summarise(shapiro_test = list(shapiro.test(median_normalised_difference)), .groups = "drop") %>% 
  mutate(
    statistic = map_dbl(shapiro_test, ~ .x$statistic),
    p_value = map_dbl(shapiro_test, ~ .x$p.value)
  )
## aud_fast: p value = 0.6562552	-> normal distribution
## aud_slow: p value =	0.1551331	-> normal distribution
## multi_fast: p value =	0.3878939	-> normal distribution
## multi_slow: p value =	0.5107778	-> normal distribution
## vis_fast: p value =	0.2649088	-> normal distribution
## vis_slow: p value = 0.5639413	-> normal distribution

comp_task_diff_distribution <- normalised_tempo %>% 
  group_by(comparison, task_type) %>% 
  summarise(shapiro_test = list(shapiro.test(median_normalised_difference)), .groups = "drop") %>% 
  mutate(
    statistic = map_dbl(shapiro_test, ~ .x$statistic),
    p_value = map_dbl(shapiro_test, ~ .x$p.value)
  )
## faster	aud: p value =	0.1551331	-> normal distribution
## faster	multi: p value =	0.5107778	-> normal distribution
## faster	vis: p value =	0.3454721	-> normal distribution
## slower	aud: p value =	0.6562552	-> normal distribution
## slower	multi: p value =	0.3878939	-> normal distribution
## slower	vis:	 p value =	0.4573564	-> normal distribution

#homogeneity of variance

##frequency
leveneTest(median_normalised_difference ~ frequency, normalised_tempo) #p value = 0.4006 -> homogeneity
##comparison of pre stim and stim tempo
leveneTest(median_normalised_difference ~ comparison, normalised_tempo) #p value = 0.3571 -> homogeneity
##task_type
leveneTest(median_normalised_difference ~ task_type, normalised_tempo) #p value = 0.09704 -> homogeneity
##condition
leveneTest(median_normalised_difference ~ condition, normalised_tempo) #p value = 0.6974 -> homogeneity
##comparison of pre stim and stim tempo * task_type
leveneTest(median_normalised_difference ~ comparison * task_type, normalised_tempo) #p value = 0.6864 -> homogeneity
```

###ANOVA -> generally significant results
```{r}
#frequency * task_type
anova1 <- aov(median_normalised_difference ~ frequency * task_type, data = normalised_tempo)
summary(anova1)

#comparison * task_type
anova2 <- aov(median_normalised_difference ~ comparison * task_type, data = normalised_tempo)
summary(anova2)

#condition
anova3 <- aov(median_normalised_difference ~ condition, data = normalised_tempo)
summary(anova3)


```
### Post-hoc testing
```{r}

```


##H2: The multimodal stimuli will influence the spontaneous motor tempo more than the unimodal stimuli. 

## Visualisations
```{r}
ggplot(normalised_tempo, aes(frequency, median_normalised_difference, color = frequency)) +
  geom_boxplot()

ggplot(normalised_tempo, aes(comparison, median_normalised_difference, color = comparison)) +
  geom_boxplot()

ggplot(clean_df_plot, aes(condition, tempo, color = event)) +
  geom_boxplot()
```

