---
title: "PerAct_Exam"
<<<<<<< Updated upstream
author: "Zofia Radwańska, Silke Swärd, Valeria Alladio"
date: "2025-01-06"
=======
author: "Zofia Radwańska"
date: "2024-12-09"
>>>>>>> Stashed changes
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
# downloading the packages 
library(pacman)
pacman::p_load(tidyverse, dplyr)
```

#PREPROCESSING OF THE DATA

##Editing of the csv's: participant id & time difference between taps
```{r}
#adding participant ids to the csv's

#folder path
data_folder <- "/Users/zofiaradwanska/Desktop/Perception and Action/Exam/Perc_Act_csv"

csv_files <- list.files(path = data_folder, pattern = "*.csv", full.names = TRUE)

#vector for participant ids
participant_ids <- paste0(seq_along(csv_files))

#loop through files to add participant id and times as columns
for (i in seq_along(csv_files)) {
  #reading the csv files
  df <- read.csv(csv_files[i])
  
  #adding a column with participant id
  df$participant_id <- participant_ids[i]
  
  #creating a column with times between events
  df$time <- c(NA, diff(df$time))
  
  #saving the files
  write.csv(df, csv_files[i], row.names = FALSE)
  print(paste("Updated and saved file:", csv_files[i]))
}

```

##Editing of the data frame
```{r}
df <- list.files(path = "/Users/zofiaradwanska/Desktop/Perception and Action/Exam/Perc_Act_csv", full.names = TRUE) %>% 
  lapply(read_csv) %>%
  bind_rows

str(df)
df$participant_id <- as.factor(df$participant_id)

#changing the variable names
df$task_type <- ifelse(df$task_type == "Metronome", "aud",
                ifelse(df$task_type == "Visual Pulsation", "vis",
                ifelse(df$task_type == "Both", "multi", df$task_type)))

df$frequency <- ifelse(df$frequency == "Slow", "slow",
                ifelse(df$frequency == "Fast", "fast", df$frequency))

df$event <- ifelse(df$event == "second_tapping_keypress", "after_stim",
            ifelse(df$event == "first_tapping_keypress", "before_stim", df$event))

#adding a condition column
df <- df %>% 
  mutate(condition = paste(task_type, frequency, sep = "_"))

#deleting the time difference for the first tap in second tapping condition (the inaccurate time recording)
df <- df %>%
  group_by(participant_id) %>% 
  mutate(
    is_first_after = event == "after_stim" & row_number() == which(event == "after_stim")[1], #marking the first after_stim tap
    time = ifelse(is_first_after, NA, time)
  ) %>%
  ungroup() %>% 
  select(-is_first_after)
```

##Data inspection
```{r}
#initial data inspection
df_dirty_summary <- df %>%
  group_by(condition, event) %>%
  filter(!event %in% c("image_description", "pulsation")) %>% 
  summarize(
    mean_tempo = mean(time, na.rm = TRUE),
    median_tempo = median(time, na.rm = TRUE),
    min_tempo = min(time, na.rm = TRUE),
    max_tempo = max(time, na.rm = TRUE),
    sd_tempo = sd(time, na.rm = TRUE),
    .groups = "drop"  # removes all standard grouping 
  )

#visual inspection
df_plot <- df %>%  #df for plotting (only the tapping times are included)
  filter(!event %in% c("pulsation", "image_description"))

#time differences against timeline for each condition
ggplot(df_plot, aes(timestamp, time, color = condition)) +
  geom_point() +
  geom_line()

#time difference against the timeline for each participant
ggplot(df_plot, aes(timestamp, time, color = participant_id)) +
  geom_point() +
  geom_line()
<<<<<<< Updated upstream

#time difference against the timeline for each participant (separately)
participant_plots <- lapply(unique(df_plot$participant_id), function(id) {
  ggplot(df_plot %>% filter(participant_id == id),
         aes(x = timestamp, y = time, color = as.factor(participant_id))) +
    geom_point() +
    geom_line() +
    theme_minimal() +
    labs(
      title = paste("Participant", id),
      x = "Timestamp",
      y = "Time Between Taps"
    ) +
    theme(legend.position = "none")
})

##individual plots
participant_plots[[19]]  # checking the plot by participant id

=======
>>>>>>> Stashed changes
```

##Data cleaning
```{r}
#removal of participant 29 -> didn't understand the instructions
data <- df %>% 
  filter(!participant_id %in% 29)

#summary of the time differences
data_sum <- data %>% 
  group_by(participant_id, event, condition) %>% 
  filter(!event %in% c("image_description")) %>% 
  summarize(
    mean_tempo = mean(time, na.rm = TRUE),
    median_tempo = median(time, na.rm = TRUE),
    min_tempo = min(time, na.rm = TRUE),
    max_tempo = max(time, na.rm = TRUE),
    sd_tempo = sd(time, na.rm = TRUE),
    .groups = "drop"  # removes all standard grouping 
  )

df_vis_fast <- data %>% 
  filter(participant_id)
  filter(condition %in% "vis_fast")
```

<<<<<<< Updated upstream
##Participant statistics 
```{r}

df <- list.files(path = "C:/Users/silke/OneDrive/Skrivbord/Aarhus Universitet/2024-2025/HT24/Perception and Action/PerAct Paper/Perc_Act_csv", full.names = TRUE) %>% 
  lapply(read_csv) %>%
  bind_rows

#mean age
mean_age = mean(df$participant_age)
print(paste("Mean age:", mean_age))

#sd age
sd_age = sd(df$participant_age)
print(paste("sd age:", sd_age))

#total number of languages 
nr_lang = length(unique(df$l1))
print(paste("Number of L1:", nr_lang))

#total number of native countries
native_country = length(unique(df$country))
print(paste("Number of nationalities:", native_country))

#new df with only ID and l1
df_w_lang <- df %>% group_by(participant_id) %>% summarise(l1 = first(l1)) 
print(df_w_lang)

#counting how many speakers there are of each language
nr_speaker <- df_w_lang %>% group_by (l1) %>% summarise(count = n())
nr_speaker

#separating those with musical training 
df_music <- df %>% filter (musical_experience == "Yes") %>% mutate(formal_education = as.numeric(formal_education))

#finding out how many had musical training
part_music <- length(unique(df_music$participant_id))
part_music

#mean of years of training 
mean_training = mean(df_music$formal_education, na.rm = TRUE)
mean_training
print(paste("Mean years of training", mean_training))

sd_train = sd(df_music$formal_education, na.rm = TRUE)
sd_train

#separating gender
df_gend <- df %>% group_by(participant_id) %>% summarise(gender = first(gender))

#counting genders 
nr_gend <- df_gend %>% group_by(gender) %>% summarise(count = n())
nr_gend
  
#new df with only ID and l1
df_w_lang <- df %>% group_by(participant_id) %>% summarise(l1 = first(l1)) 
print(df_w_lang)

#counting how many speakers there are of each language
nr_speaker <- df_w_lang %>% group_by (l1) %>% summarise(count = n())
nr_speaker
```
=======

>>>>>>> Stashed changes

#DATA ANALYSIS


